External Patch Group Prior Guided Internal Prior Learning for Real Image Denoising

### 2016-10-28 CVPR2017_Review_Unified.tex
add

Given a real noisy image, assume we can totally extract $N$ local patches from it. Similar to the external prior learning stage, for the $n$-th local patch ($n=1,...,N$), we extract its similar patches around it to form a noisy PG denoted by $\mathbf{Y}_{n} = \{\mathbf{y}_{n,1},...,\mathbf{y}_{n,M}\}$. Then the group mean of $\mathbf{Y}_{n}$, denoted by $\bm{\mu}_{n}$, is calculated and subtracted from each patch, leading to the mean subtracted PG $\mathbf{\overline{Y}}_{n}$. Each mean substracted PG is defined as $\mathbf{\overline{Y}}_{n}\triangleq \{\mathbf{\overline{y}}_{n,m}\}_{m=1}^{M}$ where $\mathbf{\overline{y}}_{n,m}=\mathbf{y}_{n,m}-\bm{\mu}_{n}$. For adaptivity, we project the PG $\mathbf{\overline{Y}}_{n}$ into its most suitable Gaussian component (subspace) of the GMM learned on external PGs. The subspace most suitable for $\mathbf{\overline{Y}}_{n}$ is selected by firstly calculating the posterior probability of "$\mathbf{\overline{Y}}_{n}$ belonging to the $k$th Gaussian component":
\vspace{-2mm}
\begin{equation}\label{equ4}\vspace{-1mm}
P(k|\mathbf{\overline{Y}}_{n})=\frac{\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{n,m}|\mathbf{0},\mathbf{\Sigma}_{k})}{\sum_{l=1}^{K}\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{n,m}|\mathbf{0},\mathbf{\Sigma}_{l})},
\end{equation}
and then choosing the component with the maximum A-posteriori (MAP) probability $\ln P(k|\mathbf{\overline{Y}}_{n})$.

@@

Assume we have assigned all internal noisy PGs $\{\mathbf{\overline{Y}}_{n}\}_{n=1}^{N}$ to their corresponding most suitable Gaussians or subspaces in $\{\mathcal{N}(\mathbf{0},\mathbf{\Sigma}_{k})\}_{k=1}^{K}$. Assume $\{\mathbf{\overline{Y}}_{k_{n}}\}_{n=1}^{N_{k}}$ are the noisy PGs assigned to the $k$-th subspace, such that $\sum_{k=1}^{K}N_{k}=N$ and $\mathbf{\overline{Y}}_{k_{n}}=[\mathbf{\overline{y}}_{k_{n},1},...,\mathbf{\overline{y}}_{k_{n},M}]$. We consider to utilize the external orthogonal dictionary $\mathbf{U}_{k}$ to guide the learning of an orthogonal dictionary $\mathbf{D}_{k}:=[\mathbf{D}_{k,e}\ \mathbf{D}_{k,i}]\in \mathbb{R}^{3p^2\times 3p^2}$. This dictionary has two parts: the external part $\mathbf{D}_{k,e}=\mathbf{U}_{k}(:,1:3p^2-r)\in\mathbb{R}^{3p^2\times (3p^2-r)}$ is directly obtained from the external dictionary in Equ. (\ref{equ3}), and the internal part $\mathbf{D}_{k,i}$ is consisted of dictionary atoms adaptively learned from the internal noisy PGs $\mathbf{Y}_{k}=\{\mathbf{\overline{Y}}_{k_{n}}\}_{n=1}^{N_{k}}$. For notation simplicity, we ignore the subspace index $k$ and denote as noisy PGs $\mathbf{Y}\triangleq\{\mathbf{\overline{Y}}_{n}\}_{n=1}^{N} = [\mathbf{\overline{y}}_{1,1},...,\mathbf{\overline{y}}_{1,M},...,\mathbf{\overline{y}}_{N,1},...,\mathbf{\overline{y}}_{N,M}]$. The learning and denoising are simultaneously performed under the weighted sparse coding framework proposed as follows:
\vspace{-2mm}
\begin{equation}\label{equ5}\vspace{-2mm}
\begin{split}
&\min_{\mathbf{D}_{i},\{\bm{\alpha}_{n,m}\}}
\sum_{n=1}^{N}\sum_{m=1}^{M}(\|\mathbf{\overline{y}}_{n,m}-\mathbf{D}\bm{\alpha}_{n,m}\|_{2}^{2}+\sum_{j=1}^{3p^{2}}\lambda_{j}|\bm{\alpha}_{n,m,j}|)
\\
&
\text{s.t.}
\quad
\mathbf{D}=[\mathbf{D}_{e}\ \mathbf{D}_{i}],\ \mathbf{D}_{i}^{\top}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{\top}\mathbf{D}_{i} = \mathbf{0},
\end{split}
\end{equation}
where $\bm{\alpha}_{n,m}$ is the sparse coefficients vector of the $m$-th patch $\mathbf{\overline{y}}_{n,m}$ of the $n$-th PG $\mathbf{\overline{Y}}_{n}$ and $\bm{\alpha}_{n,m,j}$ is the $j$-th element of $\bm{\alpha}_{n,m}$. $\lambda_{j}$ is the $j$-th regularization parameter defined as
\vspace{-2mm}
\begin{equation}\label{equ6}\vspace{-2mm}
\lambda_{j} = \lambda/(\sqrt{\mathbf{S}_{j}}+\varepsilon).
\end{equation}
We employ square roots of the singular values in $\mathbf{S}$ (please refer Equ. (\ref{equ3})) as external prior weights and add a small positive number $\varepsilon$ to avoid zero denominator. Noted that $\mathbf{D}_{e}=\emptyset$ if $r=3p^{2}$ and $\mathbf{D}_{e}=\mathbf{U}_{k}$ if $r=0$. The dictionary $\mathbf{D} = [\mathbf{D}_{e}\ \mathbf{D}_{i}]$ is orthogonal by checking that:
\vspace{-2mm}
\begin{equation}\label{equ7}\vspace{-1mm}
\mathbf{D}^{\top}\mathbf{D} = 
\left[\begin{array}{c}
\mathbf{D}_{e}^{\top}
\\
\mathbf{D}_{i}^{\top}
\end{array}\right]
[\mathbf{D}_{e}\ \mathbf{D}_{i}]
=
\left[\begin{array}{cc}
\mathbf{D}_{e}^{\top}\mathbf{D}_{e}\ \mathbf{D}_{e}^{\top}\mathbf{D}_{i}
\\
\mathbf{D}_{i}^{\top}\mathbf{D}_{e}\ \mathbf{D}_{i}^{\top}\mathbf{D}_{i}
\end{array}\right]
=
\mathbf{I}
\end{equation}

@@

We evaluate the performance of the proposed framework on denoising real noisy images. The denoising is simultaneously done with the guided internal dictionary learning (DL) process. We ignore the index $k\in\{1,...,K\}$ of subspace for notation simplicity. In the denoising stage, for each subspace, the group mean vectors $\{\bm{\mu}_{n}\}_{n=1}^{N}$ of corresponding mean substracted noisy PGs $\{\mathbf{\overline{Y}}_{n}\}_{n=1}^{N}$ are saved for reconstruction. Until now, we obtain the solutions of sparse coefficients vectors $\{\hat{\bm{\alpha}}_{n,m}^{(T-1)}\}$ in (\ref{equ9}) for $n=1,...,N;m=1,...,M$ and the orthogonal dictionary $\mathbf{D}_{(T)} = [\mathbf{D}_{e}\ \mathbf{D}_{i}^{(T)}]$ in Equ. (\ref{equ10}). Then the $m$-th latent clean patch $\hat{\mathbf{y}}_{n,m}$ in the $n$-th PG $\mathbf{Y}_{n}$ is recovered by 
\vspace{-2mm}
\begin{equation}\label{equ12}\vspace{-2mm}
\hat{\mathbf{y}}_{n,m}=\mathbf{D}_{(T)}\hat{\bm{\alpha}}_{n,m}+\bm{\mu}_{n},
\end{equation}
where $n=1,...,N;m=1,...,M$.

\begin{table}
\label{alg1}
\begin{tabular}{l}
\hline
\textbf{Alg. 1}: External Patch Group (PG) Prior Guided Internal
\\
\quad \quad \quad PG Prior Learning for Image Denoising
\\
\hline
\textbf{Input:} Noisy image $\mathbf{y}$, external PG prior GMM model
\\
\textbf{Output:} The denoised image $\hat{\mathbf{x}}$.
\\
\textbf{Initialization:} $\hat{\mathbf{x}}^{(0)}=\mathbf{y}$;
\\
\textbf{for} $Ite = 1:IteNum$ \textbf{do}
\\
1. Extracting internal PGs from $\hat{\mathbf{x}}^{(Ite-1)}$;
\\
\quad\textbf{for} each PG $\mathbf{Y}_{n}$ \textbf{do}
\\
2.\quad Calculate group mean vector $\boldsymbol{\upmu}_{n}$ and form 
\\
\quad \ \ \ mean substracted PG $\mathbf{\overline{Y}}_{n}$;
\\
3.\quad Subspace selection via Equ. (\ref{equ4});
\\
\quad\textbf{end for}
\\
\quad\textbf{for} the PGs in each Subspace \textbf{do}
\\
4.\quad External PG prior Guided Internal Orthogonal
\\
\quad \ \ \ Dictionary Learning by solving (\ref{equ5});
\\
5.\quad Recover each patch in all PGs via Equ. (\ref{equ12});
\\
\quad\textbf{end for}
\\
6. Aggregate the recovered PGs of all subspaces to form
\\
\quad the recovered image $\hat{\mathbf{x}}^{(Ite)}$;
\\
\textbf{end for}
\\
\hline
\end{tabular}
\end{table}

The latent clean image $\hat{\mathbf{x}}$ is reconstructed by aggregating all the estimated PGs. Similar to \cite{pgpd}, we perform the above denoising procedures for several iterations for better denoising outputs. But different from \cite{pgpd}, we do not employ the iterative regularization strategy \cite{osher2005iterative} since it does no help to our proposed method. The proposed denoising algorithm is summarized in Alg. 1.

>>>>

Given a real noisy image, assume we can totally extract $N$ local patches from it. Similar to the external prior learning stage, for the $n$-th local patch ($n=1,...,N$), we extract its $M$ most similar patches around it to form a noisy PG denoted by $\mathbf{Y}_{n} = \{\mathbf{y}_{n,1},...,\mathbf{y}_{n,M}\}$. Then the group mean of $\mathbf{Y}_{n}$, denoted by $\bm{\mu}_{n}$, is calculated and subtracted from each patch by $\mathbf{\overline{y}}_{n,m}=\mathbf{y}_{n,m}-\bm{\mu}_{n}$, leading to the mean subtracted PG $\mathbf{\overline{Y}}_{n}\triangleq \{\mathbf{\overline{y}}_{n,m}\}_{m=1}^{M}$. For adaptivity, we project the PG $\mathbf{\overline{Y}}_{n}$ into its most suitable Gaussian component (subspace) of the GMM learned on external PGs. The subspace most suitable for $\mathbf{\overline{Y}}_{n}$ is selected by firstly calculating the posterior probability of "$\mathbf{\overline{Y}}_{n}$ belonging to the $k$th Gaussian component":
\vspace{-2mm}
\begin{equation}\label{equ4}\vspace{-1mm}
P(k|\mathbf{\overline{Y}}_{n})=\frac{\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{n,m}|\mathbf{0},\mathbf{\Sigma}_{k})}{\sum_{l=1}^{K}\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{n,m}|\mathbf{0},\mathbf{\Sigma}_{l})}
\end{equation}
for $k=1,...,K$, and then choosing the component with the maximum A-posteriori (MAP) probability $\max_{k}P(k|\mathbf{\overline{Y}}_{n})$.

@@

Assume we have assigned all internal noisy PGs $\{\mathbf{\overline{Y}}_{n}\}_{n=1}^{N}$ to their corresponding most suitable Gaussians or subspaces in $\{\mathcal{N}(\mathbf{0},\mathbf{\Sigma}_{k})\}_{k=1}^{K}$. For the $k$-th subspace, assume the noisy PGs assigned to it are $\{\mathbf{\overline{Y}}_{k_{n}}\}_{n=1}^{N_{k}}$ such that $\mathbf{\overline{Y}}_{k_{n}}=[\mathbf{\overline{y}}_{k_{n},1},...,\mathbf{\overline{y}}_{k_{n},M}]$ and $\sum_{k=1}^{K}N_{k}=N$. We consider to utilize the external orthogonal dictionary $\mathbf{U}_{k}$ (Equ. (\ref{equ3})) to guide the learning of an orthogonal dictionary defined as $\mathbf{D}_{k}\triangleq[\mathbf{D}_{k,e}\ \mathbf{D}_{k,i}]\in \mathbb{R}^{3p^2\times 3p^2}$. This dictionary has two parts: the external part $\mathbf{D}_{k,e}=\mathbf{U}_{k}(:,1:3p^2-r)\in\mathbb{R}^{3p^2\times (3p^2-r)}$ is directly obtained from the external dictionary $\mathbf{U}_{k}$, and the internal part $\mathbf{D}_{k,i}$ is consisted of dictionary atoms adaptively learned from the internal noisy PGs $\{\mathbf{\overline{Y}}_{k_{n}}\}_{n=1}^{N_{k}}$. Therefore, the dictionary $\mathbf{D}_{k}$ can adaptively characterize the internal PGs in the $k$-th subspace. For notation simplicity, we ignore the subspace index $k$ and denote the noisy PGs assigned to each subspace as $\mathbf{Y}\triangleq\{\mathbf{\overline{Y}}_{n}\}_{n=1}^{N} = [\mathbf{\overline{y}}_{1,1},...,\mathbf{\overline{y}}_{1,M},...,\mathbf{\overline{y}}_{N,1},...,\mathbf{\overline{y}}_{N,M}]$. The learning is performed under the weighted sparse coding framework proposed as follows:
\vspace{-2mm}
\begin{equation}\label{equ5}\vspace{-2mm}
\begin{split}
&\min_{\mathbf{D}_{i},\{\bm{\alpha}_{n,m}\}}
\sum_{n=1}^{N}\sum_{m=1}^{M}(\|\mathbf{\overline{y}}_{n,m}-\mathbf{D}\bm{\alpha}_{n,m}\|_{2}^{2}+\sum_{j=1}^{3p^{2}}\lambda_{j}|\bm{\alpha}_{n,m,j}|)
\\
&
\text{s.t.}
\quad
\mathbf{D}=[\mathbf{D}_{e}\ \mathbf{D}_{i}],\ \mathbf{D}_{i}^{\top}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{\top}\mathbf{D}_{i} = \mathbf{0},
\end{split}
\end{equation}
where $\bm{\alpha}_{n,m}$ is the sparse coefficients vector of the $m$-th patch $\mathbf{\overline{y}}_{n,m}$ in the $n$-th PG $\mathbf{\overline{Y}}_{n}$ and $\bm{\alpha}_{n,m,j}$ is the $j$-th element of $\bm{\alpha}_{n,m}$. $\lambda_{j}$ is the $j$-th regularization parameter defined as
\vspace{-2mm}
\begin{equation}\label{equ6}\vspace{-2mm}
\lambda_{j} = \lambda/(\sqrt{\mathbf{S}_{j}}+\varepsilon).
\end{equation}
We employ square roots of the singular values in $\mathbf{S}$ (please refer Equ. (\ref{equ3})) as external prior weights and add a small positive number $\varepsilon$ to avoid zero denominator. Noted that $\mathbf{D}_{e}=\emptyset$ if $r=3p^{2}$ and $\mathbf{D}_{e}=\mathbf{U}_{k}$ if $r=0$. The dictionary $\mathbf{D} = [\mathbf{D}_{e}\ \mathbf{D}_{i}]$ is orthogonal by checking that:
\vspace{-2mm}
\begin{equation}\label{equ7}\vspace{-1mm}
\mathbf{D}^{\top}\mathbf{D} = 
\left[\begin{array}{c}
\mathbf{D}_{e}^{\top}
\\
\mathbf{D}_{i}^{\top}
\end{array}\right]
[\mathbf{D}_{e}\ \mathbf{D}_{i}]
=
\left[\begin{array}{cc}
\mathbf{D}_{e}^{\top}\mathbf{D}_{e}\ \mathbf{D}_{e}^{\top}\mathbf{D}_{i}
\\
\mathbf{D}_{i}^{\top}\mathbf{D}_{e}\ \mathbf{D}_{i}^{\top}\mathbf{D}_{i}
\end{array}\right]
=
\mathbf{I}
\end{equation}

@@

We evaluate the performance of the proposed framework on denoising real noisy images. The denoising is simultaneously done with the guided internal dictionary learning (DL) process. We ignore the index $k\in\{1,...,K\}$ of subspace for notation simplicity. In the denoising stage, for each subspace, the group mean vectors $\{\bm{\mu}_{n}\}_{n=1}^{N}$ of corresponding mean subtracted noisy PGs $\{\mathbf{\overline{Y}}_{n}\}_{n=1}^{N}$ are saved for reconstruction. Until now, we obtain the solutions of sparse coefficients vectors $\{\hat{\bm{\alpha}}_{n,m}^{(T-1)}\}$ in (\ref{equ9}) for $n=1,...,N;m=1,...,M$ and the orthogonal dictionary $\mathbf{D}_{(T)} = [\mathbf{D}_{e}\ \mathbf{D}_{i}^{(T)}]$ in Equ. (\ref{equ10}). Then the $m$-th latent clean patch $\hat{\mathbf{y}}_{n,m}$ in the $n$-th PG $\mathbf{Y}_{n}$ is recovered by 
\vspace{-2mm}
\begin{equation}\label{equ12}\vspace{-2mm}
\hat{\mathbf{y}}_{n,m}=\mathbf{D}_{(T)}\hat{\bm{\alpha}}_{n,m}+\bm{\mu}_{n},
\end{equation}
where $n=1,...,N;m=1,...,M$.
\begin{table}
\label{alg1}
\begin{tabular}{l}
\hline
\textbf{Alg. 1}: External Patch Group (PG) Prior Guided Internal
\\
\quad \quad \quad PG Prior Learning for Image Denoising
\\
\hline
\textbf{Input:} Noisy image $\mathbf{y}$, external PG prior GMM model
\\
\textbf{Output:} The denoised image $\hat{\mathbf{x}}$.
\\
\textbf{Initialization:} $\hat{\mathbf{x}}^{(0)}=\mathbf{y}$;
\\
\textbf{for} $Ite = 1:IteNum$ \textbf{do}
\\
1. Extracting internal PGs from $\hat{\mathbf{x}}^{(Ite-1)}$;
\\
%\textbf{Guided Internal Subspace Selection:}
%\\
\quad\textbf{for} each PG $\mathbf{Y}_{n}$ \textbf{do}
\\
2.\quad Calculate group mean vector $\boldsymbol{\upmu}_{n}$ and form 
\\
\quad \ \ \ mean subtracted PG $\mathbf{\overline{Y}}_{n}$;
\\
3.\quad Subspace selection via Equ. (\ref{equ4});
\\
\quad\textbf{end for}
\\
%\textbf{Guided Internal Orthogonal Dictionary Learning:}
%\\
\quad\textbf{for} the PGs in each Subspace \textbf{do}
\\
4.\quad External PG prior Guided Internal Orthogonal
\\
\quad \ \ \ Dictionary Learning by solving (\ref{equ5});
\\
5.\quad Recover each patch in all PGs via Equ. (\ref{equ12});
\\
\quad\textbf{end for}
\\
6. Aggregate the recovered PGs of all subspaces to form
\\
\quad the recovered image $\hat{\mathbf{x}}^{(Ite)}$;
\\
\textbf{end for}
\\
\hline
\end{tabular}
\end{table}
The latent clean image $\hat{\mathbf{x}}$ is reconstructed by aggregating all the estimated PGs. Similar to \cite{pgpd}, we perform the above denoising procedures for several iterations for better denoising outputs. The proposed denoising algorithm is summarized in Alg. 1.
